---
title: "Statistical Inference Project 1 - Part 1"
author: "Joao Clemencio"
date: "27 September 2015"
output: pdf_document
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Overview

In this project we will attempt to show that the averages of many exponential distributions follow the Central Limit Theorem, having an approximately Gaussian distibution.

# Simulations

To run the simulations the `rexp` function was used, with 40 simulations and lambda = 0.2. What follows is an example of this simulation.
```{r simulation_sample, fig.width=3, fig.height=3}
set.seed(267)
n = 40
lambda = 0.2
sample.sim = rexp(n, lambda)
summary(sample.sim)
qplot(sample.sim, binwidth = 2, main = "Sample Distribution")
```
We will now run 1000 of these simulations. The `exp.sims` variable is a matrix with 1000 rows (for each simulation group) and 40 columns (for each simulation in each group).
```{r sims}
sims = 1000
n = 40
lambda = 0.2

exp.sims = matrix(
  data = rexp(n*sims, lambda),
  nrow = 1000,
  ncol = 40,
  byrow = TRUE)
```
# Sample Mean vs Theoretical Mean
Calculating the means for each of the simulations
```{r means_plot, fig.height=3, fig.width=3}
exp.means = apply(exp.sims, 1, mean)
theo.mean = 1/lambda

c("Expected Mean" = theo.mean, "Actual Mean of Means" = mean(exp.means))

ggplot(data = data.frame(m = exp.means) %>% tbl_df, aes(x=m)) + geom_histogram() + geom_vline(xintercept = mean(exp.means), colour = "green", linetype = "dashed") + geom_vline(xintercept = theo.mean, colour = "red", linetype = "twodash")

theo.sd = 1/lambda
```
# Sample Variance vs Theoretical Variance

```{r vars_plot, fig.height=3, fig.width=3}
exp.vars = apply(exp.sims, 1, var)
theo.var = (1/lambda)^2

c("Expected Variance" = theo.var, "Actual Mean of Variances" = mean(exp.vars))

ggplot(data = data.frame(m = exp.vars) %>% tbl_df, aes(x=m)) + geom_histogram() + geom_vline(xintercept = mean(exp.vars), colour = "green", linetype = "dashed") + geom_vline(xintercept = theo.var, colour = "red", linetype = "twodash")
```

# Show that the distribution is apparently normal

```{r dens_plot, fig.height=4, fig.width=4}
# We first plot the means
hist(exp.means, breaks = n, prob = T, xlab = "Means", ylab="Density", main="Density of means")
# And then plot the distribution over it
x.fit = seq(min(exp.means), max(exp.means), length=100)
y.fit = dnorm(x.fit, mean = 1/lambda, sd = (1/lambda/sqrt(n)))
lines(x.fit, y.fit)
```

For a better view of the apparent normality, we can use a Q-Q plot.
```{r qq_plot, fig.height=4, fig.width=4}
qqnorm(exp.means)
qqline(exp.means)
```
